---
title: "stock prediction"
format: html
editor: visual
---

#### Load necessary libraries

```{r}
#| code-fold: true
library(prophet)
library(lubridate)
library(ggplot2)
library(pacman)
pacman::p_load(data.table, fixest, BatchGetSymbols, finreportr, ggplot2, lubridate)

```

::: callout-note
## Set parameters
:::

```{r}
first.date <- Sys.Date() -2500
last.date <- Sys.Date()
freq.data <- "daily"
tickers <- c("BALKRISIND.NS")
```

::: callout-note
## Collecting data

simply taking daily data from 2014-07-01 to 2022-05-05
:::

### Getting Data

```{r}
#| code-fold: true
stocks <- BatchGetSymbols(tickers=tickers,
                          first.date = first.date,
                          last.date = last.date,
                          freq.data = freq.data,
                          do.cache = FALSE,
                          thresh.bad.data = 0)
data <- stocks$df.tickers
data<- na.omit(data)
head(data)
```

```{r}
str(data)
```

::: callout-important
## dataset info

the dataset contains total `1691` observations and `10` variables
:::

```{r}
qplot(data$ref.date, data$price.close, data=data)
```

::: callout-caution
## caution

it is clearly evident that the data set is not stationary, so our next step is to make use of log transformation and convert it into stationary data
:::

#### Log transformation

```{r}
ds <- data$ref.date
y <- log(data$price.close)
df <- data.frame(ds,y)
head(df)
```

#### Stock forecasting using prophet package

```{r}
m <- prophet(df)
future <- make_future_dataframe(m, periods = 30)
forecast <- predict(m, future)
```

```{r}
f_d <- prophet(df, daily.seasonality = TRUE)
future_d <- make_future_dataframe(f_d, periods = 30)
forecast_1 <- predict(f_d, future_d)
```

### Model performance & Stock Prediction

```{r}
pred <- forecast$yhat[1:dim(df)[1]]
actual <- m$history$y
plot(actual, pred)
```

```{r}
pred_1 <- forecast_1$yhat[1:dim(df)[1]]
actual_1 <- f_d$history$y
plot(actual_1, pred_1)
```

```{r}
summary(lm(pred~actual))
```

::: callout-note
## Model 1

Adjusted R square is `98.34%` which means is a good model
:::

```{r}
summary(lm(pred_1~actual_1))
```

::: callout-note
## Model 2

basically the same
:::

### Plot forecast

```{r}
prophet_plot_components(m,forecast)
```

### predictions in R

```{r}
plot(m, forecast)
```

::: callout-tip
## interpretation

plot is showing an increasing trend for the next 30 days.

-   lets transform log values into original values based on exp function in R
:::

```{r}
tail(forecast)
forecast$yhat <- exp(forecast$yhat)
forecast$trend <- exp(forecast$trend)
forecast$trend_upper <- exp(forecast$trend_upper)
forecast$trend_lower <- exp(forecast$trend_lower)
```

::: callout-note
## predicted values

above are the predicted values in yhat
:::

::: callout-warning
## Disclaimer

for any kind of investment consult your financial advisors, i do not take any take responsibility of your loss or any trading ideas
:::
